\AorE{Wavelet Scattering Transformasies Toegepas op Walvis Vokaliserings}{Wavelet Scattering Transforms Applied to Whale Vocalisations}

### 2024-05-27 dissertation/content/p3_intro.tex
draft edit nearly complete, need to do refs

In chapter \ref*{chap:p2i}, we reviewed a generalised definition of a multi-dimensional TF decomposition $X$, which is processed by a set of invariance-enforcing filters $\Phi_i$. Wavelet scattering in 1 dimension utilising a single invariance-enforcing filter, $\phi$, as discussed in chapter \ref*{chap:p2}.

@@

Wavelet scattering filterbanks have been extended to multiple dimensions using various different wavelets and filterbank construction techniques. This chapter provides an overview of the existing filterbanks and filters and their properties. We then identify a gap in the literature, which gives rise to a generalisation of cascasded 1D scattering transforms -- separable wavelet scattering.

\section{Types of Scattering Transforms}
Existing scattering transforms are listed and described below for convenience:
\begin{enumerate}
    \item 1D scattering \citep{1dscattering1, ws}. This is the first-introduced scattering transform, which serves as a generalisation of MFCCs. It improves on MFCCs by recovering the information lost due to invariance filter (averaging) operations. Morlet wavelets are used, with cascaded scattering levels having its own number of filter per octave.
    \item 2D scattering \citep{2dscattering}. Gabor filters modified to 0-mean, are used to construct the filterbank. The filterbank consists of exponentially placed filters in one axis, which is then rotated by equally spaced angles to cover the frequency plane. This filterbank only has a single filter per octave ($Q=1$). In particular, in polar coordinates $(r, \theta)$ in the frequency domain, filters are placed exponentially along $r$, but linearly along $\theta$.
    \item Harmonic scattering \citep{harmonicscattering} (3 dimensions). The spherical harmonics are used as filters which consitute an orthogonal basis in 3 dimensions. These filters have ``lobes'' that extend in a multitude of dimensions, depending on the harmonic, in the frequency domain.
    \item 3D scattering -- extension of 2D scattering \citep{3dscattering}. The filterbank is constructed similar to 2D scattering. In spherical coordinates $(\rho, \theta, \phi)$ in the frequency domain, filters are spaced exponentially along $\rho$, and linearly along $\theta$ and $\phi$.
    \item Joint time-frequency \citep{ws_joint_tf,jointtfscattering2} scattering -- 2D filters on 1D scattering scalogram. A first level of scattering in performed in 1D to produce a scalogram ($X[k, m]$). 2D scattering is then performed on subsequent levels, with $X[k, m]$ as input. The 2D wavelets are separable - a product of 1D wavelets. This allows 1D scattering on audio to have features that are also dependent on adjascent frequency bins. It is conceptually similar to a \ac{cnn} operating on a \ac{stft} spectrogram. In the frequency domain with axes $(\omega_1, \omega_2)$, the 2D filters are constructed as the cross-product between filter sets that placed exponentially in both $\omega_1$ and $\omega_2$.
    This is the only literature considering separable filters.
\end{enumerate}

@@

>>>>

In chapter \ref*{chap:p2i}, we reviewed a generalised definition of a multi-dimensional TF decomposition $X$, which is processed by a set of invariance-enforcing filters $\Phi_i$. Wavelet scattering in 1 dimension utilises a single invariance-enforcing filter, $\phi$, as discussed in chapter \ref*{chap:p2}.

@@

Wavelet scattering filterbanks have been extended to multiple dimensions using various different wavelets and filterbank construction techniques. This chapter provides an over\-view of the existing wavelet filterbanks and filters and their properties. We then identify a gap in the literature, which gives rise to a generalisation of cascasded 1D scattering transforms -- separable wavelet scattering.

\section{Types of Scattering Transforms}
Existing scattering transforms are listed and described below for convenience:
\begin{enumerate}
    \item 1D scattering \citep{1dscattering1, ws}. This is the first-introduced scattering transform, which serves as a generalisation of MFCCs. It improves on MFCCs by recovering the information lost due to invariance filter (averaging) operations. Morlet wavelets are used, with cascaded scattering levels having its own number of filter per octave.
    \item 2D scattering \citep{2dscattering}. Gabor filters modified to have 0-mean are used to construct the filterbank. The filterbank consists of exponentially placed filters in one axis, which is then rotated by equally spaced angles to cover the frequency plane. In particular, in polar coordinates $(r, \theta)$ in the frequency domain, filters are placed exponentially along $r$ with a single filter per octave, but linearly along $\theta$.
    \item Harmonic scattering \citep{harmonicscattering} (3 dimensions). The spherical harmonics are used as filters which consitute an orthogonal basis in 3 dimensions. These filters have ``lobes'' that extend in the along various directions in the frequency domain, depending on the harmonic.
    \item 3D scattering -- extension of 2D scattering \citep{3dscattering}. The filterbank is constructed similar to 2D scattering. In spherical coordinates $(\rho, \theta, \phi)$ in the frequency domain, filters are spaced exponentially along $\rho$ with a single filter per octave, and linearly along $\theta$ and $\phi$.
    \item Joint time-frequency \citep{ws_joint_tf,jointtfscattering2} scattering -- 2D filters on 1D scattering scalogram. A first level of scattering is performed on a 1D time signal to produce a 2D scalogram ($X[k, m]$). 2D scattering is then performed on subsequent levels, with $X[k, m]$ as input. The 2D wavelets are separable - a product of 1D wavelets. This allows 1D scattering on audio to have features that are also dependent on adjacent frequency bins, which is not possible with 1D scattering only. The extracted features can therefore correlate with frequency modulations while retaining the benefits from 1D. It is conceptually similar to a \ac{cnn} operating on a \ac{stft} spectrogram. In the frequency domain with axes $(\omega_1, \omega_2)$, the 2D filters are constructed as the cross-product between filter sets that placed exponentially in both $\omega_1$ and $\omega_2$.
    This is the only literature considering separable filters.
\end{enumerate}

@@

Recently, wavelet scattering operations implemented as \acp{cnn}, referred to as a scattering network, have been of increasing interest in order to increase the effectiveness of \ac{cnn} filters \cite{2dscattering, scattering_birdsong}. Results indicate that making some aspects of scattering networks learnable yields significant improvements compared to non-learnable filters. Particularly, fixed spatial and/or time filters with learnable across-channel filters seem to obtain the best results \cite{nnphasecollapse}.

For gradient descent, it may be beneficial to perform convolutions directly, and not in the Fourier domain. Separable filters have been shown to be effective in a \ac{cnn} settings, while also allowing for much quicker convolution operations \cite{separablecnn,separablecnn2} when computed directly. Separable filters also allow for the independance of each dimension, which can be particularly useful for optimisations such as downsampling. This type of scheme is illustrated by higher-dimensional dyadic wavelet transforms such as the \ac{dwt} \cite{waveletbook}.

Downsampling schemes are trivial when the axes are independent -- a filterbank can be configured independently for each axis, where subsequent axes remain oblivious to previous operations. When a mix of time and spatial dimensions are present, existing techniques such as 2D, 3D and harmonic wavelet scattering may not be suitable, since their invariance filters cannot provide a certain amount of invariance for each dimension independently, due to their filterbank construction. Additionally, 2D and 3D scattering traditionally do not allow the user to create a filterbank with more than 1 filter per octave across $r$ and $\rho$ respectively. Signals with different time/spatial supports (i.e., a signal may be much longer in one dimension compared to another) will likely require different invariance scales across dimensions. As such, conventional higher dimensional scattering transforms may not be appropriate.

The type of scattering transform used is usually highly dependent on the signal dimen\-sionality -- there is no unifiying multi-dimensional analogy of 1D scattering as is present with dyadic wavelet transforms \citep{waveletbook}. We construct this generalisation in the chapter \ref*{chap:p3} via separable filters.

### 2024-05-27 dissertation/content/paper1/main.tex
started intro to p3
draft edit nearly complete, need to do refs

\chapter{Wavelet Transforms as a Fourier Alternative}

\articleinfo{M.W. Rademan, D.J.J Verself, J.A. Du Preez}{Ecological Informatics}{15 May 2024}{doi}{Underwater passive acoustic monitoring systems record many hours of audio data for marine research, 
making fast and reliable non-causal signal detection paramount. Such detectors assist in reducing the amount of 
labor required for signal annotations, which often contain large portions devoid of signals.\par
Cetacean vocalization detection based on spectral entropy is investigated as a means of vocalization discovery. 
Previous techniques using spectral entropy mostly consider time-frequency enhancement of the entropy measure, and 
utilize the short time Fourier transform (STFT) as its time-frequency (TF) decomposition. Spectral entropy methods 
also requires the user to set a detection threshold manually, which call for knowledge of the produced entropy measures.\par
This paper considers median filtering as a simple, effective way to provide temporal stabilization to the entropy measure, 
and considers the continuous wavelet transform (CWT) as an alternative TF decomposition. K-means clustering is used to 
determine the threshold required to accurately separate the signal/no-signal entropy measures, resulting in a  one-dimensional, 
two-class classification problem. The class means are used to perform pseudo-probabilistic soft class assignment, which is a 
useful metric in algorithmic development. The effect of median filtering, signal-to-noise ratio and the chosen TF decomposition are investigated.\par 
The accuracy and specificity measures of the proposed detection technique are simulated using a pulsed frequency modulated sweep, 
corrupted by a sample of ocean noise. The results show that median filtering is particularly effective for low signal-to-noise ratios. 
Both the STFT and CWT prove to be effective TF analyses for signal detection purposes, each presenting with different advantages and drawbacks. 
The simulated results provide insight into configuring the proposed detector, which is compared to a conventional STFT-based spectral entropy 
detector using manually annotated humpback whale (\textit{Megaptera novaeangliae}) songs recorded in False Bay, South Africa, July 2021.\par
The proposed method shows a significant improvement in detection accuracy and specificity, while also providing a more interpretable 
detection threshold setting via soft class assignment, providing a detector for use in development of adaptive algorithms.%
}

@@

		\node (cs)[right=0.3cm of Sigmoid] {\scriptsize $c_s(H[m])$};

		\draw[->] (x) -- (TF.west);
		\draw[->] (TF.east) -- (SE.west);
		\draw[->] (SE.east) -- (MF.west);
		\draw[->] (MF.east) -- (Sigmoid.west)node[above, midway]{\scriptsize $H[m]$};
		\draw[->] (Sigmoid.east) -- (cs.west);
		\draw[->] (MF.south) -- (KM.north);
		\draw[<-]  (Sigmoid.south) |- (KM.east) node[above, near end]{\scriptsize $\mu_S$, $\mu_{NS}$};

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{fbdata1}
	\caption{Audio snippet and spectrogram of a humpback song recorded in False Bay. The waveform shown is high-passed at 100 Hz, resampled and normalized. The signal annotations are shown on the spectrogram. The spectrogram is calculated before high-passing the signal.}
	\label{fig:fbdata1}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\textwidth]{real_data_hist}
	\caption{Histograms indicating the combined call length and distance between calls of the two humpback songs, for use in determining median filter window length limits.}
	\label{fig:fbhist}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.7\textwidth]{cwt_fb_example}
	\caption{An example of a CWT filterbank, shown as the L1 and L2 normalized versions , with $\gamma = 10$, $\beta=50$ and 4 voices per octave.}
	\label{fig:cwt_fb_example}
\end{figure}

@@

Given the class means $\mu_S$ (signal), $\mu_{NS}$ (no signal) determined by k-means clustering, the k-means decision boundary $\beta$ is the midpoint:
\begin{equation*}
	\beta = \mu_S + \frac{\mu_{NS} - \mu_S}{2}
\end{equation*}

In classical K-means, the assigned class $c_h(H[m])$ results in a hard output:
\begin{equation*}
	c_h(H[m]) = \begin{cases}
		1, & H[m] \leq \beta \\
		0, & H[m] > \beta
	\end{cases}
\end{equation*}

@@

To provide a soft output, a function which smoothly transitions from one class to another in the interval $H[m] \in [\mu_S, \mu_{NS}]$ is required. Since an entropy measurement which is lower than $\mu_S$ is more probable to contain signal and, conversely, an entropy measurement which is greater than $\mu_{NS}$ is more probable to be noise, the transition function $\alpha(t)$ should satisfy the following criteria:
\begin{enumerate}
	\item  $\alpha(t)$ is monotonically decreasing.
	\item Approaches hard classification in the limit: $\lim\limits_{t \rightarrow \infty} \alpha(t) = 0$ and $\lim\limits_{t \rightarrow -\infty} \alpha(t) = 1$.
	\item Equal class probability at the boundary: $\alpha(t - \beta) |_{t=\beta} = \alpha(0) =  \frac{1}{2}$ .
	\item Class assignment symmetry: $\alpha(t) = 1 - \alpha(-t)$.
\end{enumerate}

@@

The shape of the sigmoid function may be set by specifying the gain $g > 0$ as a hyper parameter. A higher gain will result in a steeper shape, thus attributing more certainty to entropy measurements closer to the class means, and vice versa for lower gains. It may also be beneficial to normalize the entropy measure such that $H[m] = \mu_S$ and $H[m] = \mu_{NS}$ map to $\pm1$. The normalization will ensure consistency in the behaviour of the hyper parameter $g$ over a range of class means. Figure \ref{fig:sigmoid} shows the effect of $g$ on the classification output.

The soft-output classification function $c_{s}(H[m])$ is therefore
\begin{equation*}
	c_{s}(H[m]) = \sigma\left(	-2g	\frac{H[m] - \beta}{\mu_{NS} - \mu_S}	\right).
\end{equation*}

To assign a specific pseudo-probability $c_{s}(H[m]) = p$ that corresponds to signal presence when $H[m] = \mu_S$, then the gain hyper parameter $g$ is calculated as
\begin{equation}
	\label{eqn:sigmoidg}
	g = \frac{1}{2} \frac{\mu_{NS} - \mu_S}{\mu_S - \beta} \ln\left(\frac{1}{p} - 1\right)
\end{equation}

@@

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.6\textwidth]{sigmoid}
	\caption{Example of the sigmoid function for soft output classification for various gains ($\mu_S=2$, $\mu_{NS}=4$).}
	\label{fig:sigmoid}
\end{figure}

@@

Given a discrete-time signal $x[n]$, and an window of odd length $M$, the median-filtered signal $y[n]$ is
\begin{equation*}
	y[n] = \text{med}(x[n - R], ..., x[0], ..., x[n + R] ),
\end{equation*}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{entropymf}
	\caption{Median filtered entropy (STFT) with $M=355$.}
	\label{fig:entropymf}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{kmeans1}
	\caption{The resulting class means from the K-means algorithm on median filtered entropy with $M=355$.}
	\label{fig:kmeans1}
\end{figure}

@@

The soft-output classification function $c_s(H[m])$ is used to determine the soft true positive (STPR) and true negative (STNR) discovery rates:
\begin{equation*}
	\text{STPR} = \frac{1}{|S_x|} \sum_{n \in S_x} c_s(H[m])
\end{equation*}
\begin{equation*}
	\text{STNR} = \frac{1}{N - |S_x|} \sum_{n \notin S_x} 1 - c_s(H[m])
\end{equation*}

This is equivalent to calculating the conventional TPR and TNR discovery rates if $c_h(H[m])$ is used instead of $c_s(H[m])$.

@@

 \begin{figure}[h!]
 	\centering
 	\includegraphics[width=0.6\textwidth]{res_snr_mf}
 	\caption{STFT and CWT soft discovery rates over a range of SNR's and median filter window lengths. The left-hand side of the figure shows the STPR and STNR of the STFT respectively. The CWT soft discovery rates are on the right-hand side.}
 	\label{fig:res_snr_mf}
 \end{figure}

@@

 \begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{res_mf_snr}
	\caption{STFT and CWT soft discovery rates over a range median filter window lengths for low SNRs. The left and right sides of the figure show the results for the STFT and CWT respectively.}
	\label{fig:res_mf_snr}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{res_classmeans}
	\caption{STFT and CWT difference in class means ($\mu_{NS} - \mu_S$) over a range of SNRs and median filter window lengths.}
	\label{fig:res_classmeans}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{res_snr_mf_L2}
	\caption{Comparing L1 and L2 CWT normalization for spectral entropy calculation. The left side of the figure shows the results without MF ($M=1$). The right side of the figure shows with MF ($M=501$). Note the large differences between L1 and L2 normalization for STNR and STPR.}
	\label{fig:res_snr_mf_L2}
\end{figure}

@@

 \begin{figure}[h!]
	\centering
	\includegraphics[width=1\textwidth]{res_realdata_snippet1}
	\caption{A snippet showing the measures produced by the entropy and the proposed soft-output signal detector.}
	\label{fig:res_realdata_snippet1}
\end{figure}

@@

 \begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\textwidth]{res_realdata_roc}
	\caption{The ROC curves comparing the proposed technique and the baseline STFT SE. Median filter window lengths of 1, 201 and 501 are shown. The sigmoid gain is set such that $p=0.95$. FPR is plotted on a log scale, which better illustrates the TPR differences for small FPR values.}
	\label{fig:res_realdata_roc}
\end{figure}

 \begin{figure}[h!]
	\centering
	\includegraphics[width=0.55\textwidth]{res_realdata_thresh}
	\caption{The thresholds used to produce the ROC curve and the TPR/FPR of the baseline SE and the proposed technique. Median filter window length of 1, 201 and 501 are shown on the right. The baseline entropy is shown on the left. }
	\label{fig:res_realdata_thresh}
\end{figure}

@@

In practice, the detector should be tuned such that the FPR is below 5\%. From figure \ref{fig:res_realdata_thresh}, this corresponds to a threshold of $c_s(H[m]) > 0.1$ for $p=0.95$. This threshold will vary depending on the choice of $p$. Also note that the CWT SE without median filtering ($M=1$) outperforms the STFT in the $\text{FPR} < 5\%$ region, indicating that it is the best TF-decomposition for similar practical applications.

@@

\changes{The addition of the k-means algorithm presents another advantage: utilizing the class-mean separation ($\mu_{NS} - \mu_S$) as metric to quickly determine whether there are points of interest within an audio segment. Furthermore, the k-means algorithm may be restricted to converge within a set number of iterations. Whether the algorithm converged indicates if there are points of interest within an audio segment. By segmenting the long recording and testing for appropriate class mean separation and k-means algorithm convergence, the proposed technique provides a fast way of discarding large portions of noise in long audio recordings. This method may be robust to changes in noise conditions, as long as the changes in noise are not experienced in a given segment of a long audio recording. The use of such a system remains to be explored.}

@@

\begin{subappendices}
	\section{K-means Implementation}
	\label{app:kmeans}
	For a known signal and noise profile, such as white noise, it is possible to estimate the expected values of the spectral entropy measure $H[m]$ given signal/no-signal conditions. However, in a real-world setting, the noise profile cannot be assumed to be a stationary process such as Gaussian white noise (GWN) or colored white noise (CWN). The variability of AMFM signals also further complicates this estimation. Setting appropriate thresholds on the entropy measure have to be done either manually, or via an automatic process.

@@

	Given the class means $\mu_S$, $\mu_{NS}$ and the entropy measure $H[m]$ of length $N$, the assignment function assigns the class with the closest mean:

	\begin{equation} 
	\label{eqn:classassign}
		w[m] = \begin{cases}
			1, & (H[m] - \mu_S)^2 \leq (H[m] - \mu_{NS})^2 \\
			0, & \text{otherwise}
		\end{cases}.
	\end{equation}

	The K-means algorithm for this application is shown below:
	\begin{enumerate}
		\item Assign initial means:
		\begin{equation*}
			\mu_S = \min\limits_{m}\left(H[m]\right);
		\end{equation*}
		\begin{equation*}
			\mu_{NS} = \max\limits_{m}\left(H[m]\right).
		\end{equation*}
		\item Perform class assignment using equation (\ref{eqn:classassign}).
		\item Update the means according to:
			\begin{equation*}
				\mu_S = \frac{\sum_{m=0}^{N-1} w[m] H[m]}{\sum_{m=0}^{N-1} w[m] };
			\end{equation*}
			\begin{equation*}
				\mu_{NS} = \frac{\sum_{m=0}^{N-1} (1-w[m]) H[m]}{\sum_{m=0}^{N-1} 1 - w[m] }.
			\end{equation*}
		\item Repeat steps 2 and 3 until the maximum change in means converges to a specified value $\epsilon$. Terminate if it exceeds a predefined number of iterations.
	\end{enumerate}

>>>>

\chapter{Wavelet Transforms as a Fourier Alternative}
\label{chap:p1}

\articleinfo{Soft-output Signal Detection for Cetacean Vocalizations using
Spectral \break Entropy, K-means Clustering and the Continuous Wavelet Transform}{Ecological Informatics}{15 May 2024}{doi}

\section*{CRediT}
\textbf{M.W. Rademan}: conceptualisation; methodology; software; formal analysis; writing -- original draft.

\textbf{D.J.J. Verseld}: data collection; writing -- review and editing; supervision.

\textbf{J.A. Du Preez}: writing -- review and editing; supervision.

\section*{Abstract}
Underwater passive acoustic monitoring systems record many hours of audio data for marine research, 
making fast and reliable non-causal signal detection paramount. Such detectors assist in reducing the amount of labor required for signal annotations, which often contain large portions devoid of signals.

Cetacean vocalization detection based on spectral entropy is investigated as a means of vocalization discovery. Previous techniques using spectral entropy mostly consider time-frequency enhancement of the entropy measure, and utilize the short time Fourier transform (STFT) as its time-frequency (TF) decomposition. Spectral entropy methods also requires the user to set a detection threshold manually, which call for knowledge of the produced entropy measures.

This paper considers median filtering as a simple, effective way to provide temporal stabilization to the entropy measure, and considers the continuous wavelet transform (CWT) as an alternative TF decomposition. K-means clustering is used to determine the threshold required to accurately separate the signal/no-signal entropy measures, resulting in a  one-dimensional, two-class classification problem. The class means are used to perform pseudo-probabilistic soft class assignment, which is a useful metric in algorithmic development. The effect of median filtering, signal-to-noise ratio and the chosen TF decomposition are investigated.

The accuracy and specificity measures of the proposed detection technique are simulated using a pulsed frequency modulated sweep, corrupted by a sample of ocean noise. The results show that median filtering is particularly effective for low signal-to-noise ratios. Both the STFT and CWT prove to be effective TF analyses for signal detection purposes, each presenting with different advantages and drawbacks. The simulated results provide insight into configuring the proposed detector, which is compared to a conventional STFT-based spectral entropy detector using manually annotated humpback whale (\textit{Megaptera novaeangliae}) songs recorded in False Bay, South Africa, July 2021.

The proposed method shows a significant improvement in detection accuracy and specificity, while also providing a more interpretable detection threshold setting via soft class assignment, providing a detector for use in development of adaptive algorithms.

@@

		\node (cs)[right=0.3cm of Sigmoid] {\scriptsize $p_\text{soft}(H[m])$};

		\draw[->] (x) -- (TF.west);
		\draw[->] (TF.east) -- (SE.west);
		\draw[->] (SE.east) -- (MF.west);
		\draw[->] (MF.east) -- (Sigmoid.west)node[above, midway]{\scriptsize $H[m]$};
		\draw[->] (Sigmoid.east) -- (cs.west);
		\draw[->] (MF.south) -- (KM.north);
		\draw[<-]  (Sigmoid.south) |- (KM.east) node[above, near end]{\scriptsize $\mu_s$, $\mu_n$};

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{fbdata1}
	\caption[Audio snippet and spectrogram of a humpback song recorded in False Bay.]{Audio snippet and spectrogram of a humpback song recorded in False Bay. The waveform shown is high-passed at 100 Hz, resampled and normalized. The signal annotations are shown on the spectrogram. The spectrogram is calculated before high-passing the signal.}
	\label{fig:fbdata1}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\textwidth]{real_data_hist}
	\caption[Humpback call length and time spread.]{Histograms indicating the combined call length and distance between calls of the two humpback songs, for use in determining median filter window length limits.}
	\label{fig:fbhist}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.7\textwidth]{cwt_fb_example}
	\caption[An example of a CWT filterbank.]{An example of a CWT filterbank, shown as the L1 and L2 normalized versions , with $\gamma = 10$, $\beta=50$ and 4 voices per octave.}
	\label{fig:cwt_fb_example}
\end{figure}

@@

Given the class means $\mu_s$ (signal), $\mu_n$ (no signal) determined by k-means clustering, the k-means decision boundary $\beta$ is the midpoint:
\begin{equation*}
	\beta = \mu_s + \frac{\mu_n - \mu_s}{2}
\end{equation*}

In classical K-means, the assigned class $p_\text{hard}(H[m])$ results in a hard output:
\begin{equation*}
	p_\text{hard}(H[m]) = \begin{cases}
		1, & H[m] \leq \beta \\
		0, & H[m] > \beta
	\end{cases}
\end{equation*}

@@

To provide a soft output, a function which smoothly transitions from one class to another in the interval $H[m] \in [\mu_s, \mu_n]$ is required. Since an entropy measurement which is lower than $\mu_s$ is more probable to contain signal and, conversely, an entropy measurement which is greater than $\mu_n$ is more probable to be noise, the transition function $\alpha(t)$ should satisfy the following criteria:
\begin{enumerate}
	\item  $\alpha(t)$ is monotonically decreasing.
	\item Approaches hard classification in the limit: $\lim\limits_{t \rightarrow \infty} \alpha(t) = 0$ and $\lim\limits_{t \rightarrow -\infty} \alpha(t) = 1$.
	\item Equal class probability at the boundary: $\alpha(t - \beta) |_{t=\beta} = \alpha(0) =  \frac{1}{2}$ .
	\item Class assignment symmetry: $\alpha(t) = 1 - \alpha(-t)$.
\end{enumerate}

@@

The shape of the sigmoid function may be set by specifying the gain $g > 0$ as a hyper parameter. A higher gain will result in a steeper shape, thus attributing more certainty to entropy measurements closer to the class means, and vice versa for lower gains. It may also be beneficial to normalize the entropy measure such that $H[m] = \mu_s$ and $H[m] = \mu_n$ map to $\pm1$. The normalization will ensure consistency in the behaviour of the hyper parameter $g$ over a range of class means. Figure \ref{fig:sigmoid} shows the effect of $g$ on the classification output.

The soft-output classification function $c_{s}(H[m])$ is therefore
\begin{equation*}
	c_{s}(H[m]) = \sigma\left(	-2g	\frac{H[m] - \beta}{\mu_n - \mu_s}	\right).
\end{equation*}

To assign a specific pseudo-probability $c_{s}(H[m]) = p$ that corresponds to signal presence when $H[m] = \mu_s$, then the gain hyper parameter $g$ is calculated as
\begin{equation}
	\label{eqn:sigmoidg}
	g = \frac{1}{2} \frac{\mu_n - \mu_s}{\mu_s - \beta} \ln\left(\frac{1}{p} - 1\right)
\end{equation}

@@

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.6\textwidth]{sigmoid}
	\caption[Example of the sigmoid function.]{Example of the sigmoid function for soft output classification for various gains ($\mu_s=2$, $\mu_n=4$).}
	\label{fig:sigmoid}
\end{figure}

@@

Given a discrete-time signal $x[n]$, and an window of odd length $M$, the median-filtered signal $y[n]$ is
\begin{equation*}
	y[n] = \text{median}(x[n - R], ..., x[0], ..., x[n + R] ),
\end{equation*}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{entropymf}
	\caption[Median filtered entropy.]{Median filtered entropy (STFT) with $M=355$.}
	\label{fig:entropymf}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{kmeans1}
	\caption[K-means class means on entropy.]{The resulting class means from the K-means algorithm on median filtered entropy with $M=355$.}
	\label{fig:kmeans1}
\end{figure}

@@

The soft-output classification function $p_\text{soft}(H[m])$ is used to determine the soft true positive (STPR) and true negative (STNR) discovery rates:
\begin{equation*}
	\text{STPR} = \frac{1}{|S_x|} \sum_{n \in S_x} p_\text{soft}(H[m])
\end{equation*}
\begin{equation*}
	\text{STNR} = \frac{1}{N - |S_x|} \sum_{n \notin S_x} 1 - p_\text{soft}(H[m])
\end{equation*}

This is equivalent to calculating the conventional TPR and TNR discovery rates if $p_\text{hard}(H[m])$ is used instead of $p_\text{soft}(H[m])$.

@@

 \begin{figure}[h!]
 	\centering
 	\includegraphics[width=0.6\textwidth]{res_snr_mf}
 	\caption[STFT and CWT soft discovery rates over SNR.]{STFT and CWT soft discovery rates over a range of SNR's and median filter window lengths. The left-hand side of the figure shows the STPR and STNR of the STFT respectively. The CWT soft discovery rates are on the right-hand side.}
 	\label{fig:res_snr_mf}
 \end{figure}

@@

 \begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{res_mf_snr}
	\caption[STFT and CWT soft discovery rates over median filter length.]{STFT and CWT soft discovery rates over a range median filter window lengths for low SNRs. The left and right sides of the figure show the results for the STFT and CWT respectively.}
	\label{fig:res_mf_snr}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{res_classmeans}
	\caption[STFT and CWT difference in class means.]{STFT and CWT difference in class means ($\mu_n - \mu_s$) over a range of SNRs and median filter window lengths.}
	\label{fig:res_classmeans}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{res_snr_mf_L2}
	\caption[Comparing L1 and L2 CWT normalization for spectral entropy calculation.]{Comparing L1 and L2 CWT normalization for spectral entropy calculation. The left side of the figure shows the results without MF ($M=1$). The right side of the figure shows with MF ($M=501$). Note the large differences between L1 and L2 normalization for STNR and STPR.}
	\label{fig:res_snr_mf_L2}
\end{figure}

@@

 \begin{figure}[h!]
	\centering
	\includegraphics[width=1\textwidth]{res_realdata_snippet1}
	\caption[A snippet showing the measures produced by the proposed entropy detector.]{A snippet showing the measures produced by the entropy and the proposed soft-output signal detector.}
	\label{fig:res_realdata_snippet1}
\end{figure}

@@

 \begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\textwidth]{res_realdata_roc}
	\caption[The ROC curves comparing the proposed technique and the baseline STFT SE.]{The ROC curves comparing the proposed technique and the baseline STFT SE. Median filter window lengths of 1, 201 and 501 are shown. The sigmoid gain is set such that $p=0.95$. FPR is plotted on a log scale, which better illustrates the TPR differences for small FPR values.}
	\label{fig:res_realdata_roc}
\end{figure}

 \begin{figure}[h!]
	\centering
	\includegraphics[width=0.55\textwidth]{res_realdata_thresh}
	\caption[The thresholds used to produce the ROC curve.]{The thresholds used to produce the ROC curve and the TPR/FPR of the baseline SE and the proposed technique. Median filter window length of 1, 201 and 501 are shown on the right. The baseline entropy is shown on the left. }
	\label{fig:res_realdata_thresh}
\end{figure}

@@

In practice, the detector should be tuned such that the FPR is below 5\%. From figure \ref{fig:res_realdata_thresh}, this corresponds to a threshold of $p_\text{soft}(H[m]) > 0.1$ for $p=0.95$. This threshold will vary depending on the choice of $p$. Also note that the CWT SE without median filtering ($M=1$) outperforms the STFT in the $\text{FPR} < 5\%$ region, indicating that it is the best TF-decomposition for similar practical applications.

@@

\changes{The addition of the k-means algorithm presents another advantage: utilizing the class-mean separation ($\mu_n - \mu_s$) as metric to quickly determine whether there are points of interest within an audio segment. Furthermore, the k-means algorithm may be restricted to converge within a set number of iterations. Whether the algorithm converged indicates if there are points of interest within an audio segment. By segmenting the long recording and testing for appropriate class mean separation and k-means algorithm convergence, the proposed technique provides a fast way of discarding large portions of noise in long audio recordings. This method may be robust to changes in noise conditions, as long as the changes in noise are not experienced in a given segment of a long audio recording. The use of such a system remains to be explored.}

@@

\pagebreak
\begin{subappendices}
	\section{K-means Implementation}
	\label{app:kmeans}
	For a known signal and noise profile, such as white noise, it is possible to estimate the expected values of the spectral entropy measure $H[m]$ given signal/no-signal conditions. However, in a real-world setting, the noise profile cannot be assumed to be a stationary process such as Gaussian white noise (GWN) or colored white noise (CWN). The variability of AMFM signals also further complicates this estimation. Setting appropriate thresholds on the entropy measure have to be done either manually, or via an automatic process.

@@

	Given the class means $\mu_s$, $\mu_n$ and the entropy measure $H[m]$ of length $N$, the assignment function assigns the class with the closest mean:

	\begin{equation} 
	\label{eqn:classassign}
		w[m] = \begin{cases}
			1, & (H[m] - \mu_s)^2 \leq (H[m] - \mu_n)^2 \\
			0, & \text{otherwise}
		\end{cases}.
	\end{equation}

	The K-means algorithm for this application is shown below:
	\begin{enumerate}
		\item Assign initial means:
		\begin{equation*}
			\mu_s = \min\limits_{m}\left(H[m]\right);
		\end{equation*}
		\begin{equation*}
			\mu_n = \max\limits_{m}\left(H[m]\right).
		\end{equation*}
		\item Perform class assignment using equation (\ref{eqn:classassign}).
		\item Update the means according to:
			\begin{equation*}
				\mu_s = \frac{\sum_{m=0}^{N-1} w[m] H[m]}{\sum_{m=0}^{N-1} w[m] };
			\end{equation*}
			\begin{equation*}
				\mu_n = \frac{\sum_{m=0}^{N-1} (1-w[m]) H[m]}{\sum_{m=0}^{N-1} 1 - w[m] }.
			\end{equation*}
		\item Repeat steps 2 and 3 until the maximum change in means converges to a specified value $\epsilon$. Terminate if it exceeds a predefined number of iterations.
	\end{enumerate}
